Local to global N input modelling

### 2025-06-10 docs/manuscript.qmd
Updates

## Production parameters

Salmon were grown for 548 days and were initialised Spring. This was on day `r unique(farm_coords$t_start[farm_coords$lat < 0])` for farms in the southern hemisphere and day `r unique(farm_coords$t_start[farm_coords$lat > 0])` for farms in the northern hemisphere. Cohorts were initialised anually, such that there was an overlap of 183 days where one cohort was nearing harvest and another was being initialised.

@@

@@

@tbl-params shows the parameter values used to model the growth of farmed Atlantic salmon.

| Parameter | Value | Units | Description | Reference | 
|:-------:|:----:|:----:|:----------------|:-------------------------|
| $\alpha$             | `r unname(params['alpha'])`      | -                        | Feeding catabolism coefficient | *Salvelinus alpinus*, *Oncorhynchus mykiss* (temps: 8–13$^{\circ}$C) [@broekhuizen_modelling_1994] | 
| $\epsilon_{P}$       | `r unname(params['epsprot'])`    | J g$^{-1}$               | Energy content of protein      | @baldan_r_2018 |
| $\epsilon_{L}$       | `r unname(params['epslip'])`     | J g$^{-1}$               | Energy content of lipid        | @baldan_r_2018 |
| $\epsilon_{C}$       | `r unname(params['epscarb'])`    | J g$^{-1}$               | Energy content of carbohydrate | @baldan_r_2018 |
| $\epsilon_{O_2}$     | `r unname(params['epsO2'])`      | J g O$_2^{-1}$           | Energy consumed by the respiration of 1 g of oxygen | @elliott_energy_1975 |
| $p_k$                | `r unname(params['pk'])`         | d$^{-1}$                 | Temperature coefficient for the fasting catabolism | @macnaughton_using_2019 |
| $k_0$                | `r unname(params['k0'])`         | -                        | Fasting catabolism at 0 $^{^\circ}$C | @macnaughton_using_2019 |
| $m$                  | `r unname(params['m'])`          | -                        | Weight exponent for anabolism | Experimentally-derived alue over a range of *Oncorhynchus*, *Perca* and *Salmo* species (temps: 8–13$^{\circ}$C) [@broekhuizen_modelling_1994] | 
| $n$                  | `r unname(params['n'])`          | -                        | Weight exponent for catabolism | @baldan_r_2018 |
| $\beta C$            | `r unname(params['betac'])`      | -                        | Shape coefficient for the feeding function | Fish size range: 2.25 $\pm$ 0.49 – 11.5 $\pm$ 5.29 g, temperature ranges: 5–24$^{\circ}$C. Coefficient for mean daily energy intake of the moderately fast growing group, 0.4398 (mean fast group), 0.2214 (mean slow group) [@jonsson_thermal_2001] | 
| $T_{max}$            | `r unname(params['Tma'])`        | $^{\circ}$C              | Maximum lethal temperature | Juvenile *Oncorhynchus tshawytscha* (19.8 $\pm$ 0.02 g, acclimation temp: 15$^{^\circ}$C, 19${\circ}$C) [@poletto_unusual_2017] |
| $T_{opt}$            | `r unname(params['Toa'])`        | $^{\circ}$C              | Optimal temperature | Fish size range: 2.25 $\pm$ 0.49 – 11.5 $\pm$ 5.29 g, temperature ranges: 5–24$^{^\circ}$C [@jonsson_thermal_2001]. Alt value: 14°C [@handeland_effect_2008], *Salmo salar* initial size: 77.0 $\pm$ 14.6 g, temps: 6, 10, 14, 18${\circ}$C | 
| $T_{min}$            | `r unname(params['Taa'])`        | $^{\circ}$C              | Lowest feeding temperature | Fish essentially stop feeding when below 2$^{\circ}$C [@vadboncoeur_lowering_2023] | 
| $\omega$             | `r unname(params['omega'])`      | g O$_2$ g$^{-1}$         | Oxygen consumption:weight loss ratio | Calibrated value from @baldan_r_2018 | 
| $a$                  | `r unname(params['a'])`          | J g tissue$^{-1}$        | Energy content of fish tissue | Alt value: 17.02350 (nls model fitted to female Atlantic salmon data extracted and transformed from @jonsson_energy_2003 [@van_tien_development_2016] |
| $k$                  | `r unname(params['k'])`          | -                        | Weight exponent for energy content | Alt value: 1.65175 (nls model fitted to female Atlantic salmon data extracted and transformed from @jonsson_energy_2003 [@van_tien_development_2016] |
| $eff$                | `r unname(params['eff'])`        | -                        | Food ingestion efficiency | Range: 95-97% ingestion efficiency [@uglem_does_2020] | 
| $\overline{W}$       | `r unname(params['meanW'])`      | g                        | Dry weight mean | Range of 100-150 given so used a mean of 125 and sd of 10 which over 1000 draws from a normal distribution delivers a minimum and maximum close to these numbers [@dempsey_estimating_2023] | 
| $\Delta{W}$          | `r unname(params['deltaW'])`     | g                        | Dry weight standard deviation | Range of 100-150 given so used a mean of 125 and sd of 10 which over 1000 draws from a normal distribution delivers a minimum and maximum close to these numbers [@dempsey_estimating_2023] | 
| $\overline{I_{max}}$ | `r unname(params['meanImax'])`   | g g$^{-1}$ fish d$^{-1}$ | Ingestion rate mean | [FAO](https://www.fao.org/fileadmin/user_upload/affris/docs/tabl10.pdf) |
| $\Delta{I_{max}}$    | `r unname(params['deltaImax'])`  | g g$^{-1}$ fish d$^{-1}$ | Ingestion rate standard deviation | [FAO](https://www.fao.org/fileadmin/user_upload/affris/docs/tabl10.pdf) |
| $\overline{F}$       | `r unname(params['overFmean'])`  | g g$^{-1}$ fish          | Overfeeding rate mean |    |
| $\Delta{F}$          | `r unname(params['overFdelta'])` | g g$^{-1}$ fish          | Overfeeding rate standard deviation |    |
| $mort$               | `r unname(params['mortmyt'])`    | d$^{-1}$                 | Natural mortality rate | @tvete_towards_2023 |
: Model growth parameters for Atlantic salmon {#tbl-params}

@@

Feeds came from [@cottrell_origins_nodate; @cottrell_global_2020].

[Plan to put full details of ingredients in a supplementary section. But possibly report these stats in Results under Uneaten feed?]{.aside}

@@

Ingredients are tracked seperately through the model with their own digestibility coefficients. Overall, proteins within the reference feed were 
`r round(100*filter(feed_summary, feed == "reference", macro == "P")$digest, 1)`% digestible, proteins within the past feed were 
`r round(100*filter(feed_summary, feed == "past", macro == "P")$digest, 1)`% digestible, and proteins within the future feed were 
`r round(100*filter(feed_summary, feed == "future", macro == "P")$digest, 1)`% digestible.
The digestibility of the other macronutrients are assumed to remain constant across all ingredients at `r round(100*filter(feed_summary, feed == "past", macro == "C")$digest, 1)`% for carbohydrates and `r round(100*filter(feed_summary, feed == "past", macro == "L")$digest, 1)`% for lipids.

@@

*To detail here:*

* How to convert feed and uneaten faeces into nutrients - just nitrogen? Carbon too? If just nitrogen, then can probably use the standard N = 6.25 x protein conversion to compare the feeds (lipids and carbohydrates have negligible protein). If doing carbon too, will need to find conversion rates for all of them, possibly splitting carbs/lipids into complex/simple, saturated/unsaturated, etc.
* Source of species vulnerability layers

@@

## Uneaten feed

>>>>

@@

<!-- Salmon were grown for 548 days and were initialised in Spring. This was on day `r unique(farm_coords$t_start[farm_coords$lat < 0])` for farms in the southern hemisphere and day `r unique(farm_coords$t_start[farm_coords$lat > 0])` for farms in the northern hemisphere. Cohorts were initialised annually, such that there was an overlap of 183 days where one cohort was nearing harvest and another was being initialised. -->

@@

| Parameter | Value | Reference |
|:------:|:------:|:-----------:|
| $\alpha$             | `r unname(params['alpha'])`      | [@broekhuizen_modelling_1994] |
| $\epsilon_{P}$       | `r unname(params['epsprot'])`    | @baldan_r_2018 |
| $\epsilon_{L}$       | `r unname(params['epslip'])`     | @baldan_r_2018 |
| $\epsilon_{C}$       | `r unname(params['epscarb'])`    | @baldan_r_2018 |
| $\epsilon_{O_2}$     | `r unname(params['epsO2'])`      | @elliott_energy_1975 |
| $p_k$                | `r unname(params['pk'])`         | @macnaughton_using_2019 |
| $k_0$                | `r unname(params['k0'])`         | @macnaughton_using_2019 |
| $m$                  | `r unname(params['m'])`          | @broekhuizen_modelling_1994 | 
| $n$                  | `r unname(params['n'])`          | @baldan_r_2018 |
| $\beta C$            | `r unname(params['betac'])`      | @jonsson_thermal_2001 |
| $T_{max}$            | `r unname(params['Tma'])`        | @poletto_unusual_2017 | 
| $T_{opt}$            | `r unname(params['Toa'])`        | @jonsson_thermal_2001 | 
| $T_{min}$            | `r unname(params['Taa'])`        | @vadboncoeur_lowering_2023 | 
| $\omega$             | `r unname(params['omega'])`      | @baldan_r_2018 |
| $a$                  | `r unname(params['a'])`          | @van_tien_development_2016 |
| $k$                  | `r unname(params['k'])`          | @van_tien_development_2016 | 
| $eff$                | `r unname(params['eff'])`        | @uglem_does_2020 | 
| $\overline{W}$       | `r unname(params['meanW'])`      | @dempsey_estimating_2023 | 
| $\Delta{W}$          | `r unname(params['deltaW'])`     | @dempsey_estimating_2023 | |
| $\overline{I_{max}}$ | `r unname(params['meanImax'])`   | @lall_nutrition_2009 | 
| $\Delta{I_{max}}$    | `r unname(params['deltaImax'])`  | @lall_nutrition_2009 | 
| $\overline{F}$       | `r unname(params['overFmean'])`  | |
| $\Delta{F}$          | `r unname(params['overFdelta'])` | |
| $mort$               | `r unname(params['mortmyt'])`    | @tvete_towards_2023 |
: Model growth parameters for Atlantic salmon (for double-checking numbers and reference list only) {#tbl-params}

@@

@@

Ingredients are tracked seperately through the model with their own digestibility coefficients. 
Overall, protein digestibility was 
`r round(100*filter(feed_summary, feed == "reference", macro == "P")$digest, 1)`%,
`r round(100*filter(feed_summary, feed == "past", macro == "P")$digest, 1)`%, and 
`r round(100*filter(feed_summary, feed == "future", macro == "P")$digest, 1)`% and carbohydrate digestibility was 
`r round(100*filter(feed_summary, feed == "reference", macro == "C")$digest, 1)`%,
`r round(100*filter(feed_summary, feed == "past", macro == "C")$digest, 1)`%, and 
`r round(100*filter(feed_summary, feed == "future", macro == "C")$digest, 1)`%
within the reference, past, and future feeds respectively. 
The digestibility of lipids was assumed to remain constant across all ingredients at `r round(100*filter(feed_summary, feed == "past", macro == "L")$digest, 1)`%.

@@

General thoughts is that a risk-assessment approach is most compelling:

* Hazard: Nutrient input. Present as a change between feeds (%?), relative to background levels (ratio?)
  * Possibly ask Camille how the quotas for Tasmanian salmon are calculated
  * Background levels are from [oceanhealthindex.org](https://knb.ecoinformatics.org/view/resource_map_doi:10.5063/F1610XPS) - double-check that it includes wastewater input AND agriculture - should include both - if they're not combined ask Ben because he has them combined somewhere
  * If this is too uncertain perhaps combine into categorical (oligotrophic/eutrophic) classification?
* Exposure: Species presence. Use IUCN/Aquamap data to estimate the % of each species range is occupied by farms
  * give each farm a 1km radius, check for overlaps
  * Gage already has the data to process the Aquamap data
* Vulnerability: species vulnerability to eutrophication + nutrient input - from @butt_trait-based_2022. Possibly also use IUCN threat levels for prioritisation?

@@

## Uneaten feed and faeces
